---
title: "R Lab 3 (Visualization with R)"
author: "Italo Rossi"
date: "2024-10-06"
output: pdf_document
---

This repository contains a set of exercises for the GreenLab Lab 4 of 2024 (11th October 2024). The exercises are designed to help students get familiar with the ggplot2 package, some of its different plot types and how customizable and flexible they are.

You can consult `ggplot2_cheat_sheet.pdf` as a reference point.

## Exercise 0: Before Plotting

Install and load ggplot2

```{r ex0_package, include=FALSE}

#install.packages('ggplot2')
library(ggplot2)
```

Load `Run_Table_TPCH.csv` as a data frame.

```{r ex0_df}

df <- read.csv('data/Run_Table_TPCH.csv')

```

For the following exercises, we will call each combination of `library` and `dataframe_size` an "**experiment"**.

## Exercise 1: Boxplot

### Task 1

Create a boxplot to visualize the distribution of `energy_usage` for each **"experiment"**. Use `facet_wrap()` to separate the experiments by library or dataframe size as you see fit.

### Task 2

Label the x-axis as **"Dataframe Size"** and the y-axis as **"Energy Usage"**

```{r ex1}

p_boxplot <- ggplot(df, aes(x = dataframe_size, y = energy_usage, fill = library)) +
  geom_boxplot() +
  facet_wrap(~library) +
  theme_minimal() +
  labs(
    title = "Energy Usage by Library and Dataframe Size",
    x = "Dataframe Size",
    y = "Energy Usage",
    fill = "Library"
    ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "lightgray"),
    strip.text = element_text(face = "bold")
    )

p_boxplot
```

## Exercise 2: Violin Plot

### Task 3

Create a violin plot of `energy_usage` for each **"experiment"**.

### Task 4

Overlay a boxplot inside the violin plot to show statistical summary. Adjust its opacity if necessary for better visibility.

### Task 5

Add `geom_jitter()` to overlay individual datapoints. Adjust the size of the points if necessary for better visibility.

```{r ex2}

p_violin <- ggplot(df, aes(x=dataframe_size, y=energy_usage, fill=library)) +
  geom_violin(trim=FALSE) +
  geom_boxplot(width=0.25, fill="white", alpha=0.5, outlier.size = 0.2) + # alpha = opacity
  geom_jitter(size=0.25) +
  facet_wrap(~library) +
  stat_summary(fun = mean, geom = "point", shape = 1, size = 2) +
  labs(title="Violin Plot with Inner Boxplot and Jitter",
  x="Dataframe Sze",
  y="Energy Usage")

p_violin
```

## Exercise 3: Boxplot + mean and standard error

### Task 6

Create a boxplot of `energy_usage` for each **"experiment"**. If you created a ggplot object containing your plot from Exercise 1, you can reuse it.

### Task 7

Use `stat_summary()` to show the mean and standard error on the boxplot.

```{r ex3}

# we can simply add stat_summary to the ggplot object created for Exercise 1 as needed
p_boxplot +
  stat_summary(
    fun = mean,
    geom = "point",
    shape = 18,
    size = 3,
    color = "white",
    position = position_dodge(width = 0.75)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.2,
    color = "white",
    position = position_dodge(width = 0.75)
  )
```

## Exercise 4: Scatter Plot

### Task 8

Create a scatter plot (`geom point()`) to show how `energy_usage` varies over `trial`.

### Task 9

Assign a unique color to each **"experiment"** and add a legend

```{r ex4}
ggplot(df, aes(x = trial, y=energy_usage, color = interaction(library,dataframe_size))) +
  geom_point(alpha = 0.75) +
  labs(
    title = "Energy Usage by Experiment Trial",
    x = "Trial",
    y = "Energy Usage",
    color = "Experiment"
  )
```

## Exercise 5: Linear Regression trend line in Scatter Plot

### Task 10

Using faceting, create a grid with 4 scatter plots (one per "experiment") showing how energy_usage correlates to execution_time, Including trend lines.

```{r ex5_10}

ggplot(df, aes(x = execution_time, y = energy_usage)) +
  geom_point(alpha = 0.75) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_grid(library ~ dataframe_size) +
  labs(
    title = "Energy Usage by Execution Time",
    x = "Execution Time",
    y = "Energy Usage"
  )
```

### Task 11

If you plot `memory_usage` or `cpu_usage` as a function of `execution_time`, can you conclude whether memory or CPU are correlated to energy in any of the **"experiments"**? (Hint: looking at task 10, is energy always correlated to time in the same direction on all experiments?)

```{r ex5_11}
ggplot(df, aes(x = execution_time, y = memory_usage)) +
  geom_point(alpha = 0.75) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_grid(library ~ dataframe_size) +
  labs(
    title = "Memory Usage by Execution Time",
    x = "Execution Time",
    y = "Memory Usage"
  )

ggplot(df, aes(x = execution_time, y = cpu_usage)) +
  geom_point(alpha = 0.75) +
  geom_smooth(method = "lm", se = TRUE) +
  facet_grid(library ~ dataframe_size) +
  labs(
    title = "CPU Usage by Execution Time",
    x = "Execution Time",
    y = "CPU Usage"
  )
```

## Exercise 6: Density Plot

### Task 12

Create a density plot for `energy_usage` for each **"experiment"**, separate them using `facet_grid()`.

```{r ex6}
ggplot(df, aes(x = energy_usage, fill = dataframe_size)) +
  geom_density(alpha = 0.5) +
  facet_grid(library ~ dataframe_size)
```

## Exercise 7: LaTeX Table

LaTeX is a typesetting system commonly used in scientific publishing.

### Task 13

Use dplyr functions to calculate the average `cpu_usage`, `memory_usage` and `energy_usage` for each "experiment".

### Task 14

Using the xtable package, format the summary table as LaTeX. You can copy the output for rendering in engines such as Overleaf or LaTeXiT.

```{r ex7}
# Load required libraries
library(xtable)
library(dplyr)

# Create summary statistics
summary_stats <- df %>%
  group_by(library, dataframe_size) %>%
  summarise(
    cpu_avg = mean(cpu_usage),
    memory_avg = mean(memory_usage),
    energy_avg = mean(energy_usage),
    .groups = 'drop'
  )

# Create and format the xtable
summary_table <- xtable(summary_stats,
                       caption = "Summary Statistics by Library and DataFrame Size",
                       digits = c(0, 0, 0, 2, 2, 2))

# Print the table (you can adjust the type of output as needed)
print(summary_table, 
      include.rownames = TRUE,
      floating = FALSE)
```
