---
title: "R Lab 3 (Visualization with R)"
author: "Italo Rossi"
date: "2024-10-06"
output: pdf_document
---

This repository contains a set of exercises for the GreenLab Lab 4 of 2024 (11th October 2024). The exercises are designed to help students get familiar with the ggplot2 package, some of its different plot types and how customizable and flexible they are.

You can consult `ggplot2_cheat_sheet.pdf` as a reference point.

## Exercise 0: Before Plotting

Install and load ggplot2

```{r ex0_package, include=FALSE}

install.packages('ggplot2')
library(ggplot2)
```

Load `Run_Table_TPCH.csv` as a data frame.

```{r ex0_df}

df <- read.csv('data/Run_Table_TPCH.csv')

```

For the following exercises, we will call each combination of `library` and `dataframe_size` an "**experiment"**.

## Exercise 1: Boxplot

### Task 1

Create a boxplot to visualize the distribution of `energy_usage` for each **"experiment"**

### Task 2

Label the x-axis as **"Dataframe Size"** and the y-axis as **"Energy Usage"**

```{r ex1}

p_boxplot <- ggplot(df, aes(x = dataframe_size, y = energy_usage, fill = library)) +
  geom_boxplot() +
  facet_wrap(~library) +
  theme_minimal() +
  labs(
    title = "Energy Usage by Library and Dataframe Size",
    x = "Dataframe Size",
    y = "Energy Usage",
    fill = "Library"
    ) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.background = element_rect(fill = "lightgray"),
    strip.text = element_text(face = "bold")
    )

p_boxplot
```

## Exercise 2: Violin Plot

### Task 3

Create a violin plot of `energy_usage` for each **"experiment"**.

### Task 4

Overlay a boxplot inside the violin plot to show statistical summary. Adjust its opacity if necessary for better visibility.

### Task 5

Add \`geom_jitter()\` to overlay individual datapoints. Adjust the size of the points if necessary for better visibility.

```{r ex2}

p_violin <- ggplot(df, aes(x=dataframe_size, y=energy_usage, fill=library)) +
  geom_violin(trim=FALSE) +
  geom_boxplot(width=0.25, fill="white", alpha=0.5, outlier.size = 0.2) + # alpha = opacity
  geom_jitter(size=0.25) +
  facet_wrap(~library) +
  stat_summary(fun = mean, geom = "point", shape = 1, size = 2) +
  labs(title="Violin Plot with Inner Boxplot and Jitter",
  x="Dataframe Sze",
  y="Energy Usage")

p_violin
```

## Exercise 3: Boxplot + mean and standard error

### Task 6

Create a boxplot of `energy_usage` for each **"experiment"**. If you created a ggplot object containing your plot from Exercise 1, you can reuse it.

### Task 7

Use `stat_summary()` to show the mean and standard error on the boxplot.

```{r ex3}

# we can simply add stat_summary to the ggplot object created for Exercise 1 as needed
p_boxplot +
  stat_summary(
    fun = mean,
    geom = "point",
    shape = 18,
    size = 3,
    color = "white",
    position = position_dodge(width = 0.75)
  ) +
  stat_summary(
    fun.data = mean_se,
    geom = "errorbar",
    width = 0.2,
    color = "white",
    position = position_dodge(width = 0.75)
  )
```

## Exercise 4: Scatter Plot

### Task 8

Create a scatter plot (`geom point()`) to show how `energy_usage` varies over `trial`.

### Task 9

Assign a unique color to each **"experiment"** and add a legend

```{r ex4}
ggplot(df, aes(x = trial, y=energy_usage, color = interaction(library,dataframe_size))) +
  geom_point(alpha = 0.75) +
  labs(
    title = "Energy Usage by Experiment Trial",
    x = "Trial",
    y = "Energy Usage",
    color = "Experiment"
  )
```

## Exercise 5: Density Plot

### Task 10

Create a density plot for `energy_usage` for each **"experiment"**.

```{r ex5}
ggplot(df, aes(x = energy_usage, fill = dataframe_size)) +
  geom_density(alpha = 0.5) +
  facet_wrap(~library)
```
